---
title: "101C BBall model"
author: "William Zhu"
date: '2019-11-24'
output: html_document
---

Loading data and required packages

```{r setup, include=FALSE}
getwd()
setwd("/Users/frankzhu/dev/101C modelling/fall-2019-stats-101c")
train = read.csv('train.csv')
test = read.csv('test.csv')

library(ggplot2)
library(dplyr)
library(randomForest)
```


Combining the test and training dataset and removing ID vectors as they don't provide statistical insights.

```{r}
dim(train)
dim(test)

test_labels = test$id
game_labels = test$gameID
test$HTWins <- NA
dim(test)
all <- rbind(train, test)
all$id <- NULL
all$gameID <- NULL

summary(all)
dim(all)
str(all)

plot(all$HTWins)
```

Splitting Variables into groups.

```{r}
#HT and VT variables
VT_vars = names(all[,grepl("VT", names(all))])
HT_vars = names(all[,grepl("HT", names(all))])

#HT and VT team variables
VT_team_vars_TS = names(all[,grepl("VT.TS", names(all))])
VT_team_vars_TA = names(all[,grepl("VT.TA", names(all))])
VT_team_vars_OTS = names(all[,grepl("VT.OTS", names(all))])
VT_team_Vars_OTA = names(all[,grepl("VT.OTA", names(all))])

HT_team_vars_TS = names(all[,grepl("HT.TS", names(all))])
HT_team_vars_TA = names(all[,grepl("HT.TA", names(all))])
HT_team_vars_OTS = names(all[,grepl("HT.OTS", names(all))])
HT_team_Vars_OTA = names(all[,grepl("HT.OTA", names(all))])

#HT and VT player variables
VT_team_vars_S1 = names(all[,grepl("VT.S1", names(all))])
VT_team_vars_S2 = names(all[,grepl("VT.S2", names(all))])
VT_team_vars_S3 = names(all[,grepl("VT.S3", names(all))])
VT_team_Vars_S4 = names(all[,grepl("VT.S4", names(all))])
VT_team_Vars_S5 = names(all[,grepl("VT.S5", names(all))])

VT_team_vars_OS1 = names(all[,grepl("VT.OS1", names(all))])
VT_team_vars_OS2 = names(all[,grepl("VT.OS2", names(all))])
VT_team_vars_OS3 = names(all[,grepl("VT.OS3", names(all))])
VT_team_Vars_OS4 = names(all[,grepl("VT.OS4", names(all))])
VT_team_Vars_OS5 = names(all[,grepl("VT.OS5", names(all))])

HT_team_vars_S1 = names(all[,grepl("HT.S1", names(all))])
HT_team_vars_S2 = names(all[,grepl("HT.S2", names(all))])
HT_team_vars_S3 = names(all[,grepl("HT.S3", names(all))])
HT_team_Vars_S4 = names(all[,grepl("HT.S4", names(all))])
HT_team_Vars_S5 = names(all[,grepl("HT.S5", names(all))])

HT_team_vars_OS1 = names(all[,grepl("HT.OS1", names(all))])
HT_team_vars_OS2 = names(all[,grepl("HT.OS2", names(all))])
HT_team_vars_OS3 = names(all[,grepl("HT.OS3", names(all))])
HT_team_Vars_OS4 = names(all[,grepl("HT.OS4", names(all))])
HT_team_Vars_OS5 = names(all[,grepl("HT.OS5", names(all))])

#PMX Variables
VT_pmx = names(all[,grepl("VT.pmx", names(all))])

HT_pmx = names(all[,grepl("HT.pmx", names(all))])

scale(all[,HT_team_vars_S1], center = T)
```


Feature Engineering

The first thing i'm going to do is create a player efficiency variable. I'm basing this metric as a simplified version of NBA player efficiency formulas by doing a sum of pts,ast,and stl per minute.

I'm going to try it on the HT.S1 variables first
```{r}

all[, HT_team_vars_S1]
```

```{r}
set.seed(2019) 
summary(all)
dim(all)
which(is.na(all), arr.ind=TRUE)
quick_RF <- randomForest(x=all[1:9520,], y=all$HTWins[1:9520], ntree=100,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")

all[1:5,  c('VT','HT')]


```



