---
title: "101C BBall model"
author: "William Zhu"
date: '2019-11-24'
output: html_document
---

Loading data and required packages

```{r setup, include=FALSE}
train <- read.csv("~/dev/101C modelling/fall-2019-stats-101c/train.csv")
test <- read.csv("~/dev/101C modelling/fall-2019-stats-101c/test.csv")

library(ggplot2)
library(dplyr)
library(randomForest)
library(class)
library(glmnet)
```


Combining the test and training dataset and removing ID vectors as they don't provide statistical insights.

```{r message=FALSE, warning=FALSE}
test_labels = test$id
test$HTWins <- NA

all <- rbind(train, test)
all$id <- NULL
all$gameID <- NULL
```

Splitting Variables into groups.

```{r}
#HT and VT variables
VT_vars = names(all[,grepl("VT", names(all))])
HT_vars = names(all[,grepl("HT", names(all))])

#HT and VT team variables
VT_team_vars_TS = names(all[,grepl("VT.TS", names(all))])
VT_team_vars_TA = names(all[,grepl("VT.TA", names(all))])
VT_team_vars_OTS = names(all[,grepl("VT.OTS", names(all))])
VT_team_vars_OTA = names(all[,grepl("VT.OTA", names(all))])

HT_team_vars_TS = names(all[,grepl("HT.TS", names(all))])
HT_team_vars_TA = names(all[,grepl("HT.TA", names(all))])
HT_team_vars_OTS = names(all[,grepl("HT.OTS", names(all))])
HT_team_vars_OTA = names(all[,grepl("HT.OTA", names(all))])

#HT and VT player variables
VT_team_vars_S1 = names(all[,grepl("VT.S1", names(all))])
VT_team_vars_S2 = names(all[,grepl("VT.S2", names(all))])
VT_team_vars_S3 = names(all[,grepl("VT.S3", names(all))])
VT_team_vars_S4 = names(all[,grepl("VT.S4", names(all))])
VT_team_vars_S5 = names(all[,grepl("VT.S5", names(all))])

VT_team_vars_OS1 = names(all[,grepl("VT.OS1", names(all))])
VT_team_vars_OS2 = names(all[,grepl("VT.OS2", names(all))])
VT_team_vars_OS3 = names(all[,grepl("VT.OS3", names(all))])
VT_team_vars_OS4 = names(all[,grepl("VT.OS4", names(all))])
VT_team_vars_OS5 = names(all[,grepl("VT.OS5", names(all))])

HT_team_vars_S1 = names(all[,grepl("HT.S1", names(all))])
HT_team_vars_S2 = names(all[,grepl("HT.S2", names(all))])
HT_team_vars_S3 = names(all[,grepl("HT.S3", names(all))])
HT_team_vars_S4 = names(all[,grepl("HT.S4", names(all))])
HT_team_vars_S5 = names(all[,grepl("HT.S5", names(all))])

HT_team_vars_OS1 = names(all[,grepl("HT.OS1", names(all))])
HT_team_vars_OS2 = names(all[,grepl("HT.OS2", names(all))])
HT_team_vars_OS3 = names(all[,grepl("HT.OS3", names(all))])
HT_team_vars_OS4 = names(all[,grepl("HT.OS4", names(all))])
HT_team_vars_OS5 = names(all[,grepl("HT.OS5", names(all))])

#PMX Variables
VT_pmx = names(all[,grepl("VT.pmx", names(all))])

HT_pmx = names(all[,grepl("HT.pmx", names(all))])
```


##Condensing the Data

In this section I'll only be using the VT and HT variables under TS and TA.
I wont be using OTS and OTA as each row already has VT.TS, HT.TS and VT.TA, HT.TA, so adding these two category of variable is in a sense reundant.

I'll be considering TS variables as offensive stats and TA variables as defensive stats for both the visiting and offensive team.


##Creating player variables

I'm going to join:
  - HT.S# player variables with VT.OS# player variables by plmin as a shared column
  - VT.S# and HT.OS# by plmin

```{r}

VT.player1 = data.frame(all[,VT_team_vars_S1], all[,HT_team_vars_OS1[-1]])
colnames(VT.player1) <- sub("HT.OS1.", "VT.S1.", colnames(VT.player1))

VT.player2 = data.frame(all[,VT_team_vars_S2], all[,HT_team_vars_OS2[-1]])
colnames(VT.player2) <- sub("HT.OS2.", "VT.S2.", colnames(VT.player2))

VT.player3 = data.frame(all[,VT_team_vars_S3], all[,HT_team_vars_OS3[-1]])
colnames(VT.player3) <- sub("HT.OS3.", "VT.S3.", colnames(VT.player3))

VT.player4 = data.frame(all[,VT_team_vars_S4], all[,HT_team_vars_OS4[-1]])
colnames(VT.player4) <- sub("HT.OS4.", "VT.S4.", colnames(VT.player4))

VT.player5 = data.frame(all[,VT_team_vars_S5], all[,HT_team_vars_OS5[-1]])
colnames(VT.player5) <- sub("HT.OS5.", "VT.S5.", colnames(VT.player5))

HT.player1 = data.frame(all[,HT_team_vars_S1], all[,VT_team_vars_OS1[-1]])
colnames(HT.player1) <- sub("VT.OS1.", "HT.S1.", colnames(HT.player1))

HT.player2 = data.frame(all[,HT_team_vars_S2], all[,VT_team_vars_OS2[-1]])
colnames(HT.player2) <- sub("VT.OS2.", "HT.S2.", colnames(HT.player2))

HT.player3 = data.frame(all[,HT_team_vars_S3], all[,VT_team_vars_OS3[-1]])
colnames(HT.player3) <- sub("VT.OS3.", "HT.S3.", colnames(HT.player3))

HT.player4 = data.frame(all[,HT_team_vars_S4], all[,VT_team_vars_OS4[-1]])
colnames(HT.player4) <- sub("VT.OS4.", "HT.S4.", colnames(HT.player4))

HT.player5 = data.frame(all[,HT_team_vars_S5], all[,VT_team_vars_OS5[-1]])
colnames(HT.player5) <- sub("VT.OS5.", "HT.S5.", colnames(HT.player5))
```


```{r}
team.offense.df = data.frame(all[,HT_team_vars_TS[6:13]],
                             all[,VT_team_vars_TS[6:13]])


team.defense.df = data.frame(all[,HT_team_vars_TA[6:13]],
                             all[,VT_team_vars_TA[6:13]])


player.var = data.frame(VT.player1,
                        VT.player2,
                        VT.player3,
                        VT.player4,
                        VT.player5,
                        HT.player1,
                        HT.player2,
                        HT.player3,
                        HT.player4,
                        HT.player5)

other.var  = data.frame(all[,c('HTWins', 'VT', 'HT', 'VTleague', 'HTleague', 'VTcumRest', 'HTcumRest')])
                     
```

DataFrame library

```{r}
#Full Condensed Dataframe
full.df = data.frame(other.var,
                     team.offense.df,
                     team.defense.df,
                     player.var
                     )

train.full.df = full.df[1:9520,]
test.full.df = full.df[9521:11168,]
```

Variable Selection by Random Forest

```{r fig1, fig.height = 5, fig.width = 5}
knitr::opts_chunk$set(fig.width=5, fig.height=5) 

set.seed(2019) 
quick_RF.full <- randomForest(x=train.full.df[,-1], y=train.full.df$HTWins, ntree=1000, importance=TRUE)
imp_RF.full <- importance(quick_RF.full)
imp_DF.full <- data.frame(Variables = row.names(imp_RF.full), MSE = imp_RF.full[,1])
imp_DF.full <- imp_DF.full[order(imp_DF.full$MSE, decreasing = TRUE),]

ggplot(imp_DF.full[1:41,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none") + geom_hline(yintercept = 4, colour = "red") + ggtitle("Variable Importance by Random Forest")

sel.var.full = imp_DF.full[imp_DF.full$MSE > 4, ]

RF.full.train = train.full.df[,c(as.vector(sel.var.full$Variables))]
RF.full.train = data.frame(RF.full.train, train.full.df$HTWins)

RF.full.test = test.full.df[,c(as.vector(sel.var.full$Variables))]
```

##Lasso model

```{r}
x <- model.matrix(RF.full.train$train.full.df.HTWins ~ ., RF.full.train)
y <- ifelse(RF.full.train$train.full.df.HTWins=="Yes",1,0)

RF.full.test$HTWins <- sample(c("Yes", "No"),nrow(RF.full.test),replace = TRUE, prob = c(0.5,0.5))
x_test <- model.matrix(HTWins ~. ,RF.full.test)
```

```{r}
set.seed(1)
cv.out_lasso <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure = "mse")

plot(cv.out_lasso, xlab = "Lambda for Lasso")

lambda_min <- cv.out_lasso$lambda.min
#best value of lambda
lambda_1se <- cv.out_lasso$lambda.1se
#regression coefficients
coef(cv.out_lasso,s=lambda_min)

lasso_prob <- predict(cv.out_lasso,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
lasso_predict <- rep("No",nrow(RF.full.test))
lasso_predict[lasso_prob>=.5] <- "Yes"
table(lasso_predict)
test_prediction_df_lasso <- data.frame(test_labels, lasso_predict)

colnames(test_prediction_df_lasso) <- c("id","HTWins")
```

##Ridge Model

```{r}
set.seed(2)
cv.out_ridge <- cv.glmnet(x,y,alpha=0,family="binomial",type.measure = "mse")

plot(cv.out_ridge, xlab = "Lambda for Ridge")

lambda_min <- cv.out_ridge$lambda.min
#best value of lambda
lambda_1se <- cv.out_ridge$lambda.1se
#regression coefficients
coef(cv.out_ridge,s=lambda_min)

#predict class, type=”class”
ridge_prob <- predict(cv.out_ridge,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
ridge_predict <- rep("No",nrow(RF.full.test))
ridge_predict[ridge_prob>=.5] <- "Yes"

test_prediction_df_ridge <- data.frame(test_labels, ridge_predict)

colnames(test_prediction_df_ridge) <- c("id","HTWins")
```