---
title: "101C BBall model"
author: "William Zhu"
date: '2019-11-24'
output: html_document
---

Loading data and required packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
getwd()
setwd("/Users/frankzhu/dev/101C modelling/fall-2019-stats-101c")
train = read.csv('train.csv')
test = read.csv('test.csv')

library(ggplot2)
library(dplyr)
library(randomForest)
library(class)
library(glmnet)
```


Combining the test and training dataset and removing ID vectors as they don't provide statistical insights.

```{r message=FALSE, warning=FALSE}
dim(train)
dim(test)

test_labels = test$id
game_labels = test$gameID
test$HTWins <- NA
dim(test)
all <- rbind(train, test)
all$id <- NULL
all$gameID <- NULL

summary(all)
dim(all)
str(all)
```

Splitting Variables into groups.

```{r}
#HT and VT variables
VT_vars = names(all[,grepl("VT", names(all))])
HT_vars = names(all[,grepl("HT", names(all))])

#HT and VT team variables
VT_team_vars_TS = names(all[,grepl("VT.TS", names(all))])
VT_team_vars_TA = names(all[,grepl("VT.TA", names(all))])
VT_team_vars_OTS = names(all[,grepl("VT.OTS", names(all))])
VT_team_vars_OTA = names(all[,grepl("VT.OTA", names(all))])

HT_team_vars_TS = names(all[,grepl("HT.TS", names(all))])
HT_team_vars_TA = names(all[,grepl("HT.TA", names(all))])
HT_team_vars_OTS = names(all[,grepl("HT.OTS", names(all))])
HT_team_vars_OTA = names(all[,grepl("HT.OTA", names(all))])

#HT and VT player variables
VT_team_vars_S1 = names(all[,grepl("VT.S1", names(all))])
VT_team_vars_S2 = names(all[,grepl("VT.S2", names(all))])
VT_team_vars_S3 = names(all[,grepl("VT.S3", names(all))])
VT_team_vars_S4 = names(all[,grepl("VT.S4", names(all))])
VT_team_vars_S5 = names(all[,grepl("VT.S5", names(all))])

VT_team_vars_OS1 = names(all[,grepl("VT.OS1", names(all))])
VT_team_vars_OS2 = names(all[,grepl("VT.OS2", names(all))])
VT_team_vars_OS3 = names(all[,grepl("VT.OS3", names(all))])
VT_team_vars_OS4 = names(all[,grepl("VT.OS4", names(all))])
VT_team_vars_OS5 = names(all[,grepl("VT.OS5", names(all))])

HT_team_vars_S1 = names(all[,grepl("HT.S1", names(all))])
HT_team_vars_S2 = names(all[,grepl("HT.S2", names(all))])
HT_team_vars_S3 = names(all[,grepl("HT.S3", names(all))])
HT_team_vars_S4 = names(all[,grepl("HT.S4", names(all))])
HT_team_vars_S5 = names(all[,grepl("HT.S5", names(all))])

HT_team_vars_OS1 = names(all[,grepl("HT.OS1", names(all))])
HT_team_vars_OS2 = names(all[,grepl("HT.OS2", names(all))])
HT_team_vars_OS3 = names(all[,grepl("HT.OS3", names(all))])
HT_team_vars_OS4 = names(all[,grepl("HT.OS4", names(all))])
HT_team_vars_OS5 = names(all[,grepl("HT.OS5", names(all))])

#PMX Variables
VT_pmx = names(all[,grepl("VT.pmx", names(all))])

HT_pmx = names(all[,grepl("HT.pmx", names(all))])
names(all[,grepl("ft", names(all))])
```


Feature Engineering

In this section I'll only be using the VT and HT variables under TS and TA.
I wont be using OTS and OTA as each row already has VT.TS, HT.TS and VT.TA, HT.TA, so adding these two category of variable is in a sense reundant.

So in summary for feature engineering, I'll be using TS variables as offensive stats and TA variables as defensive stats for both the visiting and offensive team.


From Literature Review:
---
According to the statistics in a study by Wilson (1948), nearly all
basketball games were won by the team that secured a higher
percentage of rebounds, both offensively and defensively. A few
years later, however, Galatin (1954) found that two of the most
important criteria in winning or losing basketball games on the
high school, college, and professional levels were the number of
field goals scored and the number of field goals attempted
---

---
Through detailed analysis the variables with high influence on game effectiveness were selected for final procedures. It has been proven that a limited number of factors, mostly offensive, determines sports performance in the NBA. The most critical indicators are: Win%, Offensive EFF, 3rd Quarter PPG, Win% CG, Avg Fauls and Avg Steals.
---

The paper: 

AU  - Zimmermann, Albrecht
AU  - Moorthy, Sruthi
AU  - Shi, Zifan
PY  - 2013/10/14
T1  - Predicting college basketball match outcomes using machine learning techniques: some results and lessons learned

outlines multiple statistics in section 2 that I will engineer with the given variables:
  - Possessions
  - Offensive Efficiency
  - Defensive Efficiency
  - Effective field goal percentage
  - Turnover percentage
  - Offensive Rebound Percentage
  - Free throw rate
  - Win Percent (%) ***
  - Adjusted Offensive Efficiency ***
  - Adjusted Defensive Efficiency ***
  - Win Probability ***
  
  - Player Offensive Rebound Rate ***
  - Player Defensive Rebound Rate ***
  - Assist/Turnover/Steal/Block Rate ***

*** stats still need to be made

The defensive stats will be calculated with Team Allowed variables
The offensive stats will be calculated with Team Scored variables

```{r}
#Offensive Possesions
attach(all)
all <- all %>%  mutate(HT.TS.possesions = 0.96 * (HT.TS.fga - HT.TS.oreb - HT.TS.to + (0.475*HT.TS.fta)), 
                       VT.TS.possesions = 0.96 * (VT.TS.fga - VT.TS.oreb - VT.TS.to + (0.475*VT.TS.fta)))

all[,"HT.TS.fga"]
#Defensive Possessions
all <- all %>%  mutate(HT.TA.possesions = 0.96 * (HT.TA.fga - HT.TA.oreb - HT.TA.to +   (0.475*HT.TA.fta)), 
                       VT.TA.possesions = 0.96 * (VT.TA.fga - VT.TA.oreb - VT.TA.to + (0.475*VT.TA.fta)))

#Offensive Efficiency
all <- all %>%  mutate(HT.off_eff = HT.TS.pts *100 / HT.TS.possesions, 
                       VT.off_eff = VT.TS.pts *100 / VT.TS.possesions)

#Defensive Efficiency
all <- all %>%  mutate(HT.def_eff = HT.TA.pts *100 / HT.TA.possesions, 
                       VT.def_eff = VT.TA.pts *100 / VT.TA.possesions)

#Effective Offensive Field Goal Percentage
all <- all %>%  mutate(HT.TS.efg_percent = (HT.TS.fgm + 0.5*3*HT.TS.tpm)/HT.TS.fga,
                       VT.TS.efg_percent = (VT.TS.fgm + 0.5*3*VT.TS.tpm)/VT.TS.fga)

#Turnover Percentage
all <- all %>%  mutate(HT.TS.to_percent = HT.TS.to/HT.TS.possesions,
                       VT.TS.to_percent = VT.TS.to/VT.TS.possesions)

#Offensive Rebound Percentage
all <- all %>%  mutate(HT.TS.oreb_percent = HT.TS.oreb/(HT.TS.oreb+VT.TS.dreb),
                       VT.TS.oreb_percent = VT.TS.oreb/(VT.TS.oreb+HT.TS.dreb))

#Free Throw Rate
all <- all %>%  mutate(HT.TS.ftr = HT.TS.fta/HT.TS.fga,
                       VT.TS.ftr = VT.TS.fta/VT.TS.fga)

#Offensive Field Goal Conversion
all <- all %>%  mutate(VT.TS.fg_percentage = VT.TS.fgm/VT.TS.fga, 
                       HT.TS.fg_percentage = HT.TS.fgm/HT.TS.fga)

#Field Goals Allowed Percentage
all <- all %>%  mutate(VT.TA.fg_percentage = VT.TA.fgm/VT.TS.fga, 
                       HT.TA.fg_percentage = HT.TA.fgm/HT.TA.fga)

#Three Point Conversion
all <- all %>% mutate(VT.TS.tp_percentage = VT.TS.tpm/VT.TS.tpa ,
                      HT.TS.tp_percentage = HT.TS.tpm/HT.TS.tpa) 

#Three Point Allowed Percentage
all <- all %>%  mutate(VT.TA.tp_percentage = VT.TA.tpm/VT.TS.tpa, 
                       HT.TA.tp_percentage = HT.TA.tpm/HT.TA.tpa)

```

Creating player variables

I'm first going to join:
  - HT.S# player variables with VT.OS# player variables by plmin as a shared column
  - VT.S# and HT.OS# by plmin

```{r}
#Visiting Team Players

VT.player1 = data.frame(all[,VT_team_vars_S1], all[,HT_team_vars_OS1[-1]])
colnames(VT.player1) <- sub("HT.OS1.", "VT.S1.", colnames(VT.player1))

VT.player2 = data.frame(all[,VT_team_vars_S2], all[,HT_team_vars_OS2[-1]])
colnames(VT.player2) <- sub("HT.OS2.", "VT.S2.", colnames(VT.player2))

VT.player3 = data.frame(all[,VT_team_vars_S3], all[,HT_team_vars_OS3[-1]])
colnames(VT.player3) <- sub("HT.OS3.", "VT.S3.", colnames(VT.player3))

VT.player4 = data.frame(all[,VT_team_vars_S4], all[,HT_team_vars_OS4[-1]])
colnames(VT.player4) <- sub("HT.OS4.", "VT.S4.", colnames(VT.player4))

VT.player5 = data.frame(all[,VT_team_vars_S5], all[,HT_team_vars_OS5[-1]])
colnames(VT.player5) <- sub("HT.OS5.", "VT.S5.", colnames(VT.player5))

HT.player1 = data.frame(all[,HT_team_vars_S1], all[,VT_team_vars_OS1[-1]])
colnames(HT.player1) <- sub("VT.OS1.", "HT.S1.", colnames(HT.player1))

HT.player2 = data.frame(all[,HT_team_vars_S2], all[,VT_team_vars_OS2[-1]])
colnames(HT.player2) <- sub("VT.OS2.", "HT.S2.", colnames(HT.player2))

HT.player3 = data.frame(all[,HT_team_vars_S3], all[,VT_team_vars_OS3[-1]])
colnames(HT.player3) <- sub("VT.OS3.", "HT.S3.", colnames(HT.player3))

HT.player4 = data.frame(all[,HT_team_vars_S4], all[,VT_team_vars_OS4[-1]])
colnames(HT.player4) <- sub("VT.OS4.", "HT.S4.", colnames(HT.player4))

HT.player5 = data.frame(all[,HT_team_vars_S5], all[,VT_team_vars_OS5[-1]])
colnames(HT.player5) <- sub("VT.OS5.", "HT.S5.", colnames(HT.player5))
```


Adding in these variables:

Offensive Rebounds:
  - For Team Scored variables
  - For Team Allowed variables

Defensive Rebounds:
  - For Team Scored variables
  - For Team Allowed variables
  
Assists:
  - For Team Scored variables
  - For Team Allowed variables

Steals:
  - For Team Scored variables
  - For Team Allowed variables

Turnovers:
  - For Team Scored variables
  - For Team Allowed variables

Personal Fouls:
  - For Team Scored variables
  - For Team Allowed variables
  
Block to Turnovers:
  - For Team Scored variables
  - For Team Allowed variables
  
Total Points:
  - For Team Scored variables
  - For Team Allowed variables
  
Player Plus Minus (a general measure of how effective player is)
  - For Visiting Team Players 1-5
  - For Home Team Players 1-5

Offensive and Defensive difficulty of Past Games:
  - For Visiting Team
  - For Home Team

```{r}
team.offense.df <- data.frame(all[,HT_team_vars_TS[6:13]],
                              all[,VT_team_vars_TS[6:13]],
                              all[,'HT.off_eff'],
                              all[,'VT.off_eff'],
                              all[,'HT.TS.possesions'],
                              all[,'VT.TS.possesions'],
                              all[,'HT.TS.efg_percent'],
                              all[,'VT.TS.efg_percent'],
                              all[,'HT.TS.to_percent'],
                              all[,'VT.TS.to_percent'],
                              all[,'HT.TS.oreb_percent'],
                              all[,'VT.TS.oreb_percent'],
                              all[,'HT.TS.ftr'],
                              all[,'VT.TS.ftr'],
                              all[,'HT.TS.fg_percentage'],
                              all[,'VT.TS.fg_percentage'],
                              all[,'HT.TS.tp_percentage'],
                              all[,'VT.TS.tp_percentage'])

team.defense.df <- data.frame(all[,HT_team_vars_TA[6:13]],
                              all[,VT_team_vars_TA[6:13]],
                              all[,'HT.def_eff'],
                              all[,'VT.def_eff'],
                              all[,'HT.TA.possesions'],
                              all[,'VT.TA.possesions'],
                              all[,'HT.TA.fg_percentage'],
                              all[,'VT.TA.fg_percentage'],
                              all[,'HT.TA.tp_percentage'],
                              all[,'VT.TA.tp_percentage'])


player.var = data.frame(VT.player1,
                        VT.player2,
                        VT.player3,
                        VT.player4,
                        VT.player5,
                        HT.player1,
                        HT.player2,
                        HT.player3,
                        HT.player4,
                        HT.player5)

other.var  = data.frame(all[,'HTWins'],
                        all[,'VT'],
                        all[,'HT'],
                        all[,'VTleague'],
                        all[,'HTleague'],
                        all[,'VTcumRest'],
                        all[,'HTcumRest']
                        )
```

DataFrame library

```{r}
#Full Condensed Dataframe
full.df = data.frame(other.var,
                     team.offense.df,
                     team.defense.df,
                     player.var
                     )

colnames(full.df) <- sub("all....", "", colnames(full.df))

train.full.df = full.df[1:9520,]
test.full.df = full.df[9521:11168,]


#Feature Engineered Dataframe

FE.df = data.frame(all[,'HTWins'],
                       team.offense.df[,-c(1:16)],
                       team.defense.df[,-c(1:16)]
                       )
colnames(FE.df) <- sub("all....", "", colnames(FE.df))

train.FE.df = FE.df[1:9520,]
test.FE.df = FE.df[9521:11168,]

```





Variable Selection by Random Forest

```{r}
set.seed(2019) 
quick_RF.full <- randomForest(x=train.full.df[,-1], y=train.full.df$HTWins, ntree=100, importance=TRUE)
imp_RF.full <- importance(quick_RF.full)
imp_DF.full <- data.frame(Variables = row.names(imp_RF.full), MSE = imp_RF.full[,1])
imp_DF.full <- imp_DF.full[order(imp_DF.full$MSE, decreasing = TRUE),]

ggplot(imp_DF.full[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")

sel.var.full = imp_DF.full[imp_DF.full$MSE >2, ]
as.vector(sel.var.full$Variables)

RF.full.train = train.full.df[,c(as.vector(sel.var.full$Variables))]
RF.full.train = data.frame(RF.full.train, train.full.df[,'HTWins..'])

RF.full.test = test.full.df[,c(as.vector(sel.var.full$Variables))]



####
quick_RF.FE <- randomForest(x=train.FE.df[,-1], y=train.FE.df$HTWins, ntree=100, importance=TRUE)
imp_RF.FE <- importance(quick_RF.FE)
imp_DF.FE <- data.frame(Variables = row.names(imp_RF.FE), MSE = imp_RF.FE[,1])
imp_DF.FE <- imp_DF.FE[order(imp_DF.FE$MSE, decreasing = TRUE),]

ggplot(imp_DF.FE[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")

sel.var.FE = imp_DF.FE[imp_DF.FE$MSE >0, ]
as.vector(sel.var.FE$Variables)

RF.FE.train = train.FE.df[,c(as.vector(sel.var.FE$Variables))]
RF.FE.train = data.frame(RF.FE.train, train.FE.df[,'HTWins..'])

RF.FE.test = test.FE.df[,c(as.vector(sel.var.FE$Variables))]
```


```{r}
names(RF.full.train)
RF.full.test
colnames(RF.full.train) <- sub("train.full.df....", "", colnames(RF.full.train))



RF.FE.train
colnames(RF.FE.train) <- sub("train.FE.df....", "", colnames(RF.FE.train))

RF.FE.test
```
Step-Wise Regression
```{r}
base.mod <- glm(HTWins ~ 1 , data= new.var.train, family = "binomial")  # base intercept only model
all.mod <- glm(HTWins ~ . , data= new.var.train, family = "binomial") # full model with all predictors
stepMod <- step(base.mod, scope = list(lower = base.mod, upper = all.mod), direction = "both", trace = 0, steps = 1000)  # perform step-wise algorithm
?step
shortlistedVars <- names(unlist(stepMod[[1]])) # get the shortlisted variable.
shortlistedVars <- shortlistedVars[!shortlistedVars %in% "(Intercept)"]  # remove intercept 
print(shortlistedVars)
```


```{r}
x <- model.matrix(HTWins....~.,RF.full.train)
#convert class to numerical variable
y <- ifelse(RF.full.train$HTWins....=="Yes",1,0)

#perform grid search to find optimal value of lambda
#family= binomial => logistic regression, alpha=1 => lasso
# check docs to explore other type.measure options

cv.out_ridge <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure = "mse")

plot(cv.out_ridge)

lambda_min <- cv.out_ridge$lambda.min
#best value of lambda
lambda_1se <- cv.out_ridge$lambda.1se
#regression coefficients
coef(cv.out_ridge,s=lambda_min)




RF.full.test$HTWins <- sample(c(1,0),nrow(RF.full.test),replace = TRUE, prob = c(0.5,0.5))
x_test <- model.matrix(HTWins ~. ,RF.full.test)


#predict class, type=”class”
ridge_prob <- predict(cv.out_ridge,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
ridge_predict <- rep("No",nrow(RF.full.test))
ridge_predict[ridge_prob>=.5] <- "Yes"
table(ridge_predict)
class(test_labels)
length(RF.full.test$VT..)
test_prediction_df_ridge <- data.frame(test$id, ridge_predict)

colnames(test_prediction_df_ridge) <- c("id","HTWins")
test_prediction_df_ridge$id = as.character(test_prediction_df_ridge$id)
class(test_prediction_df_ridge$id)

write.csv(test_prediction_df_ridge,file = 'ridge.RF2.full.csv', row.names = F) ### Scored .59 on Kaggle
```


```{r}
x <- model.matrix(HTWins....~.,RF.FE.train)
#convert class to numerical variable
y <- ifelse(RF.FE.train$HTWins....=="Yes",1,0)

#perform grid search to find optimal value of lambda
#family= binomial => logistic regression, alpha=1 => lasso
# check docs to explore other type.measure options

cv.out_ridge <- cv.glmnet(x,y,alpha=0,family="binomial",type.measure = "mse")

plot(cv.out_ridge)

lambda_min <- cv.out_ridge$lambda.min
#best value of lambda
lambda_1se <- cv.out_ridge$lambda.1se
#regression coefficients
coef(cv.out_ridge,s=lambda_min)




RF.FE.test$HTWins <- sample(c(1,0),nrow(RF.FE.test ),replace = TRUE, prob = c(0.5,0.5))
x_test <- model.matrix(HTWins ~. ,RF.FE.test)


#predict class, type=”class”
ridge_prob <- predict(cv.out_ridge,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
ridge_predict <- rep("No",nrow(RF.full.test))
ridge_predict[ridge_prob>=.5] <- "Yes"
table(ridge_predict)
class(test_labels)
test_prediction_df_ridge <- data.frame(test$id, ridge_predict)

colnames(test_prediction_df_ridge) <- c("id","HTWins")
test_prediction_df_ridge$id = as.character(test_prediction_df_ridge$id)
class(test_prediction_df_ridge$id)

write.csv(test_prediction_df_ridge,file = 'ridge.RF0.FE.csv', row.names = F) ### Scored .59 on Kaggle
```


