---
title: "Stats 101C_Final Project"
author: "Nishanth Shetty"
date: "November 21, 2019"
output: html_document
---

```{r}
library(readr)
Basketball_train <- read_csv("C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Datasets/Final Project/train.csv")

Basketball_test <- read_csv("C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Datasets/Final Project/test.csv")

attach(Basketball_train)

dim(Basketball_train)
dim(Basketball_test)

## Checking for NA's
sapply(Basketball_train, function(x) sum(is.na(x)))

## Are there repeat Game ID's
sum(duplicated(Basketball_train$gameID))

```
ccc:
(not all are included)
min Minutes
fgm Field Goals Made
fga Field Goals Attempted
tpm 3pt Shots Made
tpa 3pt Shots Attempted
fta Free Throws Attempted
ftm Free Throws Made
oreb Offensive Rebounds
dreb Defensive Rebounds
treb Team Rebounds
ast Assists
stl Steals
blk Blocked Shots
to Turn Overs
pf Personal Fouls
pts Total Points

## Random Forest
```{r}
set.seed(1)

which(is.na(Basketball_train),arr.ind = TRUE)

library(randomForest)

RF <- randomForest(x= Basketball_train[,-3],y = as.factor(Basketball_train$HTWins),ntree=100, importance = TRUE)

randomForest(factor(HTWins) ~ . , data= Basketball_train,importance = TRUE)

```

```{r}

Basketball_train <- Basketball_train %>%  mutate(VT.TS.fg.percentage = VT.TS.fgm/VT.TS.fga, HT.TS.fg.percentage = HT.TS.fgm/HT.TS.fga)

Basketball_train <- Basketball_train %>%  mutate(VT.TS.tp.percentage = VT.TS.tpm/VT.TS.tpa ,HT.TS.tp.percentage = HT.TS.tpm/HT.TS.tpa) 

# How many duplicated Game ID's
sum(duplicated(Basketball_train$gameID))

Basketball_test <- Basketball_test %>%  mutate(VT.TS.fg.percentage = VT.TS.fgm/VT.TS.fga, HT.TS.fg.percentage = HT.TS.fgm/HT.TS.fga)

Basketball_test <- Basketball_test %>%  mutate(VT.TS.tp.percentage = VT.TS.tpm/VT.TS.tpa ,HT.TS.tp.percentage = HT.TS.tpm/HT.TS.tpa) 

```

ccc:

fg.percentage % Field Goals Made **
tp.percentage % 3 Points Made ** 
fta Free Throws Attempted **
ftm Free Throws Made **
oreb Offensive Rebounds
dreb Defensive Rebounds
treb Team Rebounds ** 
ast Assists **
stl Steals **
blk Blocked Shots **
to Turn Overs 
pf Personal Fouls ** 
pts Total Points **


############################## 11.22 LOGISTIC REGRESSION ##############################

```{r}

summary(Basketball_train)

names(Basketball_train)

logistic_basketball <- glm(factor(HTWins) ~ VTcumRest + HTcumRest+ VT.TS.fga + VT.TS.tpa + VT.OTA.fta + HT.OTA.fga + HT.TA.pts + VT.TA.pts, data= Basketball_train, family = "binomial")

probability_predictions_training <- predict(logistic_basketball,newdata = Basketball_train,type = "response") # Predicting on training data 
predictions_training <-  ifelse(probability_predictions_training > 0.5, "Yes", "No")
mean(predictions_training != Basketball_train$HTWins) ## Training Error Rate

probability_predictions_test <- predict(logistic_basketball,newdata = Basketball_test,type = "response") #Predicting on testing data 
predictions_test <-  ifelse(probability_predictions_test > 0.5, "Yes", "No")

library(dplyr)
test_prediction_df_logistic <- cbind(Basketball_test$id,predictions_test) %>% data.frame()
colnames(test_prediction_df_logistic) <- c("id","HTWins")

write.csv(test_prediction_df_logistic,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-22_Logistic.csv') 
```


############################## 11.22 KNN ##############################
```{r}

library(class)
normalize <- function(x){
  (x -min(x))/(max(x)-min(x))
}

# Normalizing Training and Testing Data
train_X_normalized <- lapply(Basketball_train[,11:30],normalize) %>% as.data.frame()
test_X_normalized <- lapply(Basketball_test[,11:30],normalize) %>% as.data.frame()

knn_predictions <-  knn(train_X_normalized, test_X_normalized, cl= Basketball_train$HTWins, k = 5)
unique(knn_predictions)

test_prediction_df_knn <- cbind(Basketball_test$id,as.character(knn_predictions)) %>% data.frame()
colnames(test_prediction_df_knn) <- c("id","HTWins")

write.csv(test_prediction_df_knn,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-23_knn.csv') 
```


Most significant predictors from Literature: 

VT.TS.fg.percentage
HT.TS.fg.percentage
VT.TS.tp.percentage
HT.TS.tp.percentage
VT.TS.pts
HT.TS.pts
VT.TS.pf
HT.TS.pf
VT.TS.ast
HT.TS.ast
VT.TS.dreb
HT.TS.dreb
VT.TS.fta
HT.TS.fta
VT.TS.stl
HT.TS.stl

## Regsubsets to find the best subset
```{r}
library(leaps)

names(Basketball_train)

nr <- regsubsets(VT.TA.to ~ HTcumRest + VT.TS.fgm+VT.TS.fga+VT.TS.tpm+VT.TS.tpa, data = Basketball_train, method = "exhaustive")

nish_rgs <- regsubsets(as.factor(HTWins) ~ VT.TS.fg.percentage + HT.TS.fg.percentage
+VT.TS.tp.percentage
+HT.TS.tp.percentage
+VT.TS.pts
+HT.TS.pts
+VT.TS.pf
+HT.TS.pf
+VT.TS.ast
+HT.TS.ast
+VT.TS.dreb
+HT.TS.dreb
+VT.TS.fta
+HT.TS.fta
+VT.TS.stl
+HT.TS.stl, data = Basketball_train, method = "exhaustive")

dim(Basketball_test)
nish_rgs_summary <- summary(nish_rgs)
plot(1:8,nish_rgs_summary$bic,type = "b")

which.min(nish_rgs_summary$bic)


library(corrplot)
corrplot(cor(Weekly[,-9]), method = "number")

corrplot(cor(Basketball_train[,c("VT.TS.tp.percentage","VT.TS.fg.percentage","HT.TS.fg.percentage","VT.TS.tp.percentage","HT.TS.tp.percentage","VT.TS.pts","HT.TS.pts","VT.TS.pf","HT.TS.pf","VT.TS.ast","HT.TS.ast","VT.TS.dreb","HT.TS.dreb","VT.TS.fta","HT.TS.fta"
,"VT.TS.stl","HT.TS.stl")]), method = "number")

```
* The best subset is the full model 

############################## 11.24 LOGISTIC REGRESSION ##############################

```{r}
logistic_basketball_11.24 <- glm(factor(HTWins) ~ VT.TS.fg.percentage + HT.TS.fg.percentage
+VT.TS.tp.percentage
+HT.TS.tp.percentage
+VT.TS.pts
+HT.TS.pts
+VT.TS.pf
+HT.TS.pf
+VT.TS.ast
+HT.TS.ast
+VT.TS.dreb
+HT.TS.dreb
+VT.TS.fta
+HT.TS.fta
+VT.TS.stl
+HT.TS.stl, data= Basketball_train, family = "binomial")

probability_predictions_training <- predict(logistic_basketball_11.24,newdata = Basketball_train,type = "response") # Predicting on training data 
predictions_training <-  ifelse(probability_predictions_training > 0.5, "Yes", "No")
mean(predictions_training != Basketball_train$HTWins) ## Training Error Rate

probability_predictions_test <- predict(logistic_basketball_11.24,newdata = Basketball_test,type = "response") #Predicting on testing data 
predictions_test <-  ifelse(probability_predictions_test > 0.5, "Yes", "No")

## How many Y's and N's
table(predictions_test)

test_prediction_df_logistic <- cbind(Basketball_test$id,predictions_test) %>% data.frame()
colnames(test_prediction_df_logistic) <- c("id","HTWins")

write.csv(test_prediction_df_logistic,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-24_Logistic.csv') 
```

############################## 11.22 KNN ##############################
```{r}

library(class)
normalize <- function(x){
  (x -min(x))/(max(x)-min(x))
}

# Normalizing Training and Testing Data
train_X_normalized <- lapply(Basketball_train[,c("VT.TS.tp.percentage","VT.TS.fg.percentage","HT.TS.fg.percentage","VT.TS.tp.percentage","HT.TS.tp.percentage","VT.TS.pts","HT.TS.pts","VT.TS.pf","HT.TS.pf","VT.TS.ast","HT.TS.ast","VT.TS.dreb","HT.TS.dreb","VT.TS.fta","HT.TS.fta"
,"VT.TS.stl","HT.TS.stl")],normalize) %>% as.data.frame()

test_X_normalized <- lapply(Basketball_test[,c("VT.TS.tp.percentage","VT.TS.fg.percentage","HT.TS.fg.percentage","VT.TS.tp.percentage","HT.TS.tp.percentage","VT.TS.pts","HT.TS.pts","VT.TS.pf","HT.TS.pf","VT.TS.ast","HT.TS.ast","VT.TS.dreb","HT.TS.dreb","VT.TS.fta","HT.TS.fta"
,"VT.TS.stl","HT.TS.stl")],normalize) %>% as.data.frame()

knn_predictions <-  knn(train_X_normalized, test_X_normalized, cl= Basketball_train$HTWins, k = 20)
unique(knn_predictions)

## How many Y's and N's
table(knn_predictions)

test_prediction_df_knn <- cbind(Basketball_test$id,as.character(knn_predictions)) %>% data.frame()
colnames(test_prediction_df_knn) <- c("id","HTWins")

write.csv(test_prediction_df_knn,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-24_knn.csv') 

knn_predictions_5 <-  knn(train_X_normalized, test_X_normalized, cl= Basketball_train$HTWins, k = 5)

test_prediction_df_knn <- cbind(Basketball_test$id,as.character(knn_predictions_5)) %>% data.frame()
colnames(test_prediction_df_knn) <- c("id","HTWins")

write.csv(test_prediction_df_knn,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-24_knn=5.csv') ### Scored .59 on Kaggle

```

```{r}
knn_predictions_5 <-  knn(train_X_normalized, train_X_normalized, cl= Basketball_train$HTWins, k = 5)
## Error Rate with k=5
mean(knn_predictions_5 == Basketball_train$HTWins)

knn_predictions_20 <-  knn(train_X_normalized, train_X_normalized, cl= Basketball_train$HTWins, k = 20)
## Error Rate with k=20
mean(knn_predictions_20 == Basketball_train$HTWins)

knn_predictions_100 <-  knn(train_X_normalized, train_X_normalized, cl= Basketball_train$HTWins, k = 100)
## Error Rate with k=100
mean(knn_predictions_100 == Basketball_train$HTWins)

knn_predictions_200 <-  knn(train_X_normalized, train_X_normalized, cl= Basketball_train$HTWins, k = 200)
## Error Rate with k=200
mean(knn_predictions_200 == Basketball_train$HTWins)

```

** K performs better for lower K's

############################## 11.25 Simple Check of HT score and VT Score ##############################
```{r}
karan_rocks <-  ifelse(Basketball_train$HT.TS.pts/Basketball_train$VT.TS.pts >= 1, "Yes","No")

karan_rocks2 <-  ifelse(Basketball_train$VT.OTS.pts/Basketball_train$VT.TS.pts >1, "Yes","No")

mean(karan_rocks == Basketball_train$HTWins)

mean(karan_rocks2 == Basketball_train$HTWins)

Basketball_train[1:10,c("HTWins","VT.TS.pts","HT.TS.pts")]

```


############################## 11.25 CV Lasso  ##############################

```{r}
library(glmnet)
#convert training data to matrix format
x <- model.matrix(HTWins~.,Basketball_train)
#convert class to numerical variable
y <- ifelse(Basketball_train$HTWins=="Yes",1,0)

#perform grid search to find optimal value of lambda
#family= binomial => logistic regression, alpha=1 => lasso
# check docs to explore other type.measure options
cv.out <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure = "mse")

#plot result
plot(cv.out)

lambda_min <- cv.out$lambda.min
#best value of lambda
lambda_1se <- cv.out$lambda.1se
#regression coefficients
coef(cv.out,s=lambda_min)

Basketball_test$HTWins <- sample(c(1,0),nrow(Basketball_test),replace = TRUE, prob = c(0.5,0.5))

x_test <- model.matrix(HTWins ~. ,Basketball_test)
#predict class, type=”class”
lasso_prob <- predict(cv.out,newx = x_test,s=lambda_min,type="response")
#translate probabilities to predictions
lasso_predict <- rep("No",nrow(Basketball_test))
lasso_predict[lasso_prob>.5] <- "Yes"

table(lasso_predict)

test_prediction_df_lasso <- cbind(Basketball_test$id,as.character(lasso_predict)) %>% data.frame()
colnames(test_prediction_df_lasso) <- c("id","HTWins")

write.csv(test_prediction_df_lasso,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-25_Lasso.csv') ### Scored .671 on Kaggle


```


############################## 11.25 CV Ridge  ##############################

```{r}

cv.out_ridge <- cv.glmnet(x,y,alpha=0,family="binomial",type.measure = "mse")

#plot result
plot(cv.out_ridge)

lambda_min <- cv.out_ridge$lambda.min
#best value of lambda
lambda_1se <- cv.out_ridge$lambda.1se
#regression coefficients
coef(cv.out_ridge,s=lambda_min)

#predict class, type=”class”
ridge_prob <- predict(cv.out_ridge,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
ridge_predict <- rep("No",nrow(Basketball_test))
ridge_predict[ridge_prob>=.5] <- "Yes"

table(ridge_predict)

test_prediction_df_ridge <- cbind(Basketball_test$id,as.character(ridge_predict)) %>% data.frame()
colnames(test_prediction_df_ridge) <- c("id","HTWins")

write.csv(test_prediction_df_ridge,'C:/Users/Nishanth Shetty/Desktop/Recovered Data/UCLA lecture materials for all subjects/Stats 101C/Kaggle/Submissions/Stats 101C_Nishanth Shetty_11-25_Ridge.csv') ### Scored .59 on Kaggle


```
